Mining Interaction Log Data in a Creative Arts Practice using Macro Gestures and Transition Matrices

### 2015-09-20 flux-entropy-paper/flux-entropy-paper.tex
cut out some commented material at the end
minor edits and bib refs.

%condiser the following? .............

@@

%   %In this note we describe a data archive for an iPad
% musical-ensemble performance practice which has included %concerts,
% demonstrations and formal HCI studies of groups of musicians over more
% than two years. Our corpus contains data for %150 performances,
% rehearsals, and demonstrations, involving some 100 different musicians
% and several different iPad apps. %At the lowest level, data for each
% musician in a particular performance was archived as Apple iOS
% callback methods for touching and swiping. Subsequent data analysis
% has been undertaken by first classifying these data into higher level
% gesture types and then describing the transitions between these
% gestures for each artist as a function of time. We %find that a
% transition-matrix measure, which we call ``flux'', can be %used to
% distinguish different performance types and contexts, as well as
% different phases (beginning, middle and end) of all of the
% performances in our corpus.

@@

\section{General comments}

@@

%Additionally, a wider dissemination of such log data will better support the replication of findings in HCI research [ref:Wilson, Max L., et al. "RepliCHI-CHI should be replicating and validating results more: discuss." CHI'11 Extended Abstracts on Human Factors in Computing Systems. ACM, 2011.]

@@

NOTE: FORGET ABOUT THE DEMONSTRATIONS AND THE DATA COLLECTION PARTS OF THE CORPUS. THEY JUST MAKE THE STORY VAGUE. YOU DON'T INCLUDE THEM IN YOUR ANALYSIS AT ALL.

NOTE: NEED TO MODIFY ALL OF THESE GRAPHS, TABLES AND FIGURES.

@@

%TODO Mention Manovich transmutation stuff

% \section{Modelling Performer Ratings with Flux and Entropy}

% \begin{figure}
%   \centering
%   \includegraphics[width=\linewidth]{figures/stat-response-fit.pdf}
%   \caption{Transition matrix statistics (flux above and entropy below) vs Likert responses for
%     questions 3, 6 and 7, all of which exhibited significant
%     relationships under a proportional odds logistic regression model.
%     A least-squares linear model is shown with a dotted line.
%     \label{fig:stat-response-fit}}
% \end{figure}

% One final question to explore is how these flux and entropy measures
% in the interaction data relate to the performers' perceptions of their
% performance?

% To answer this question, we construct models with the flux and entropy
% measures as independent variables and the Likert responses (1 to 5) as
% the dependent variable. These measures exhibited clear relationships
% for three of the survey questions (Questions 3, 6 and 7), which are
% plotted in Figure~\ref{fig:stat-response-fit} where the least-squares
% line of best fit has a positive gradient for both flux and entropy.

% Determining the strengths of these positive relationships is
% difficult, both due to the problem of treating ordinal Likert
% responses as numeric\cite{Gardner:2007fj} and also
% due to the unknown distribution of the entropy and flux statistics. As
% a result, we use a conservative proportional odds logistic regression
% (POLR)~(using the \texttt{polr} function in \texttt{R}, from the MASS
% package \cite{Venables:2002qv}) to treat the Likert responses as an
% ordered factor response.

% The POLR results indicate a significant ($p<0.05$) relationship
% between the user-reported Likert response and the flux statistic for
% Question 3 ($t = 2.77, df = 4, p = 0.025$), Question 6
% ($t = 2.41, df = 4, p = 0.037$), and Question 7
% ($t = 2.49, df = 4, p = 0.034$). The same questions displayed a
% significant relationship for the entropy statistic: Question 3
% ($t = 2.93, df = 4, p = 0.022$), Question 6
% ($t = 2.69, df = 4, p = 0.028$), and Question 7
% ($t = 2.89, df = 4, p = 0.022$). These relationships are shown in
% Figure~\ref{fig:stat-response-fit}.

% It is notable that these questions (3, 6, and 7) all pertain to the
% \emph{perceived effect of the agent} on the performers. As the agent
% uses the flux measure to trigger new-idea messages, it appears that
% performers appreciated its intervention more in particularly creative
% performances where flux and entropy were high.

\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{references}

>>>>

%consider the following? .............

@@

%   %In this note we describe a data archive for an iPad
% musical-ensemble performance practice which has included concerts,
% demonstrations and formal HCI studies of groups of musicians over more
% than two years. Our corpus contains data for 150 performances,
% rehearsals, and demonstrations, involving some 100 different musicians
% and several different iPad apps. At the lowest level, data for each
% musician in a particular performance was archived as Apple iOS
% callback methods for touching and swiping. Subsequent data analysis
% has been undertaken by first classifying these data into higher level
% gesture types and then describing the transitions between these
% gestures for each artist as a function of time. We find that a
% transition-matrix measure, which we call ``flux'', can be used to
% distinguish different performance types and contexts, as well as
% different phases (beginning, middle and end) of all of the
% performances in our corpus.

@@

%\section{General comments}

@@

% Additionally, a wider dissemination of such log data will better
% support the replication of findings in HCI
% research~\cite{Wilson:2011ve}

@@

% TODO: NOTE: FORGET ABOUT THE DEMONSTRATIONS AND THE DATA COLLECTION
% PARTS OF THE CORPUS. THEY JUST MAKE THE STORY VAGUE. YOU DON'T
% INCLUDE THEM IN YOUR ANALYSIS AT ALL.

% TODO: NOTE: NEED TO MODIFY ALL OF THESE GRAPHS, TABLES AND FIGURES.

@@

%TODO Mention Manovich transmutation stuff
The macro-gestures that these processes reveal could also be used as
the basis for new works of new media art~\cite{Manovich:2002ly}.

\pagebreak
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{references}
